---
title: "STAT 545 Assignment B-1"
output: github_document
---
STAT 545 @ UBC
Assignment B-1: Making a function
Total Points: 100

This assignment covers making a function in R, documenting it, and testing it.

# Loading packages

```{r}
suppressPackageStartupMessages(library(datateachr)) # <- contains the data you picked!
suppressPackageStartupMessages(library(tidyverse))
```

# Assignment Exercises 

## **Exercise 1: Make a Function (25 points)**
*In this exercise, you’ll be making a function and fortifying it. The function need not be complicated. The function need not be “serious”, but shouldn’t be nonsense.*

## **Exercise 2: Document your Function (20 points)**
*In the same code chunk where you made your function, document the function using roxygen2 tags.*

One of the chunks of code that I that repeated frequently for the tidying process in Milestone 2 of the mini data analysis was the unnesting, grouping, and factoring of a variable. It would be great to make this a function to not have to copy, paste, and modify the whole chunk over and over if a lot of variables have nested observations.


```{r}
#' @title A Function to Unnest and Factor a Variable in a Tibble
#'
#' @description 'unnest_and_factor' takes a designated variable from a designated dataframe and helps to tidy
#' the data by making sure that if there are multiple values included in a single observation with a standardized
#' separator, these values are unnested--separated--and then factored for sorting, counting, and other grouping purposes.
#'
#' @param df The name of the stored tibble you want to work with - called 'df' for dataframe based on common stackoverflow usage
#' @param var The target variable from that tibble - called 'var' as short for variable
#' @param sep The separator string that the function should use to identify nested values; include quotations for proper usage - called 'sep' to mirror dplyr language
#'
#'@return The function returns the unnested tibble.
#'
#'@examples
#' unnest_and_factor(steam_games, game_details,",")
#' unnest_and_factor(apt_buildings, parking_type, ",")
#' unnest_and_factor(apt_buildings, bike_parking, " and ")
#'
#' @references
#' Syntax suggested on StackOverflow by LyzandeR
#' link: \url{https://stackoverflow.com/questions/48062213/dplyr-using-column-names-as-function-arguments}
#' More information can aso be found in the book 'Tidy Evaluation' by Lionel Henry and Hadley Wickham
#' link: \url{https://tidyeval.tidyverse.org/dplyr.html}

unnest_and_factor <- function(df,var,sep) { #where df is the dataset, var is the variable to be unnested and sep is the separator used in the observations of the nested variable
    check <- dplyr::summarise(df, is.character({{var}})) #checking if the class of the variable is a character
  if (!check[[1]])
    warning("The variable you have selected to unnest is not a character. It will be converted to a character as part of the evaluation of this function.") #warning that the function converts the existing non-character variable, if applicable, to a character in order to work
   unnested_factored <- df %>%
    mutate({{var}}:=strsplit(as.character({{var}}), sep))%>%  #separating the observations on the given separator 
    unnest({{var}}) %>% #unnesting the separated observations
    mutate({{var}}:=as.factor({{var}})) #factoring the observations for grouping capabilities
   return(unnested_factored) #returning the unnested and factored tibble
}
```


## **Exercise 3: Include examples (15 points)**
*Demonstrate the usage of your function with a few examples. Use one or more new code chunks, describing what you’re doing.*

### Example 1

Here is an unnesting that I had tried in my STAT545A assignment that places each detail about a game is its own observation. To show what has changed, after unnesting the variable using the function and getting the return of the entire tibble, I asked r to print the head of the original and tidied tibble and selected just the unnested variable and the id variable.

```{r}
#using function with `steam_games` dataset variable `game_details`

unnest_and_factor(steam_games, game_details,",")

details<-unnest_and_factor(steam_games, game_details,",") #storing the function evaluating `game_details`

head(steam_games)%>%
  select(id, game_details) #the head of the original tibble's `id` and `game_details` columns

head(details)%>%
  select(id, game_details) #the head of the unnested and factored tibble's `id` and `game_details` columns
```

### Example 2

I was also interested in testing if and how the function would work for other datasets and not just the one that I had successfully unnested before. I took a look at the other datasets in the `datateachr` package and saw that `apt_buildings` also had a nested column in the `parking_type` column, so I tried the function with that variable.  To show what has changed, after unnesting the variable using the function and getting the return of the entire tibble, I asked r to print the head of the original and tidied tibble and selected just the unnested variable and the id variable.

```{r}
#using function with `apt_buildings` dataset variable `parking_type`
unnest_and_factor(apt_buildings, parking_type, ",")

parking<-unnest_and_factor(apt_buildings, parking_type, ",") #storing the function evaluating `parking_type`


head(apt_buildings)%>%
  select(id, parking_type) #the head of the original tibble's `id` and `parking_type` columns

head(parking)%>%
  select(id, parking_type) #the head of the unnested and factored tibble's `id` and `parking_type` columns
```


### Example 3

Another column with two pieces of information in `apt_buildings`is the `bike_parking` variable. I thought this might be a good variable to try using the function on because the separator is not a comma, but the word "and." This shows the function can be used with any separator the user sets, not just a single character. To show what has changed, after unnesting the variable using the function and getting the return of the entire tibble, I asked r to print the head of the original and tidied tibble and selected just the unnested variable and the id variable.

```{r}
#using function with `apt_buildings` dataset variable `bike_parking`
unnest_and_factor(apt_buildings, bike_parking, " and ")

bike<-unnest_and_factor(apt_buildings, bike_parking, " and ") #storing the function evaluating `bike_parking`

head(apt_buildings)%>%
  select(id, bike_parking) #the head of the original tibble's `id` and `bike_parking` columns

head(bike)%>%
  select(id, bike_parking) #the head of the unnested and factored tibble's `id` and `bike_parking` columns
```

## **Exercise 4: Test the Function (25 points)**

*Running examples is a good way of checking by-eye whether your function is working as expected. But, having a formal “yes or no” check is useful when you move on to other parts of your analysis.*

*Write formal tests for your function. You should use at least three non-redundant uses of an expect_() function from the testthat package, and they should be contained in a test_that() function (or more than one). They should all pass.*

### Test 1

I first decided to test to make sure that the function would produce the same tibble as if I had done the unnesting manually. To do this, I created a test that would check that the class for the languages had been converted to being "factor" and that the overall tibble was the same as the manually performed version.

```{r}
#testing that using the function and doing the specific unnesting action produce the same results and that the unnested column is factored
#using the `steam_games` dataset and the `languages` variable, which has NAs

#count NAS
steam_games %>%
    count(is.na(languages))

#unnesting manually
unnested_languages <- steam_games %>%
  mutate(languages=strsplit(as.character(languages), ",")) %>%
  unnest(languages)%>%
  mutate(languages=as.factor(languages))

#unnesting using the custom function
function_languages <- unnest_and_factor(steam_games,languages,",")

#testing that the two unnested tibbles match
testthat::test_that("factorandmatched", {
testthat::expect_s3_class(function_languages$languages, "factor") #testing that the variable `languages` has the s3 class "factor"
testthat::expect_identical(unnested_languages, function_languages)}) #testing that the output is the same as if the unnesting had been done manually
```


### Test 2

I then created a test to make sure that the unnesting was occurring and that it wasn't somehow creating new or differently named columns. To do this, I tested to make sure that the number of rows in the new tibble was greater than the original and that the number of columns and the names of the columns had remained the same.
```{r}
#testing that the unnesting is occurring by asking r to check whether the new tibble has a higher number of rows compared to the original tibble but has the same number of columns with the same names.
#using the `apt_buildings` dataset and the `facilities_available` variable, which 

#checking NA count
apt_buildings %>%
count(is.na(facilities_available))

#unnesting manually
unnested_facilities <- apt_buildings %>%
  mutate(facilities_available=strsplit(as.character(facilities_available), "/")) %>%
  unnest(facilities_available)%>%
  mutate(facilities_available=as.factor(facilities_available))

#unnesting using the custom function
function_facilities <- unnest_and_factor(apt_buildings,facilities_available,"/")

#testing that the number of rows has increased from the original tibble but that the number and names of columns has stayed the same

testthat::test_that("rowcolcount", {
  testthat::expect_gt(nrow(function_facilities), nrow(apt_buildings)) #inequality of row count
  testthat::expect_equal(ncol(function_facilities), ncol(apt_buildings)) #matching column count
  testthat::expect_equal(colnames(function_facilities),colnames(apt_buildings)) #matching column names
                    })

```


### Test 3

Lastly, I wanted to test to make sure that the warning was working by attempting to unnest a column that was not classed as "character." To do so, I tested the `steam_games` dataset variable `id`, which is classed as a double.
```{r}
#testing that evaluating the `id` variable of `steam_games` would generate a warning.
testthat::test_that("warning", testthat::expect_warning(unnest_and_factor(steam_games, id, ",")))
```

Powered by the Academic theme for Hugo.